Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs

### 2017-12-11 redaction/Presentations/SeminaireBETA_Nancy_Oct2017/slides_BETA_v1.tex
présentation Nancy

\begin{frame}
\frametitle{Empirical specification (2)}
\textbf{The estimation strategy} \vspace{0.1cm}
\begin{itemize}
\item Make some simplifying assumptions to get the estimated equation \vspace{0.1cm}
\item About the specification of both additive and multiplicative costs  \vspace{0.1cm}
\begin{itemize}
\item[(1)] Separability between the origin country $i$ and the product $k$ dimensions (ss in Irarrazabal et al., 2015)\vspace{0.1cm}
\item[(2)] All products $k$ in a 3-digit sector $s$ share the same structure of costs \vspace{0.1cm}
\item[$\Leftrightarrow$] Write $t_{ik}$ and $\tau_{ik}$ as:
\footnotesize
\begin{equation}
 \tau_{ik} = \tau_{i} \times \tau_{s(k)}, \qquad t_{ik} = t_{i} + t_{s(k)} \label{eq:specifTC}
 \end{equation}
 \normalsize
\end{itemize}
\item About the specification of the error term \vspace{0.1cm}
\begin{itemize}
\item[-] Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} -1>0$, the error term should be always positive and multiplicative \vspace{0.1cm}
\item[$\Rightarrow$] The equation of interest becomes:
\footnotesize
$$\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{s(k)} -1+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})$$
\normalsize
\begin{itemize}
\item[-] With $\epsilon_{ik}$ following a normal law centered on 0
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[label = slide_result1]
\begin{itemize}
\item \textbf{International transport costs: A sizeable additive component}  \vspace{0.2cm}
\begin{itemize}
\item[-] 48.2\% of total costs in average for vessel, 42.3\% for air  \vspace{0.1cm}
\item[-] A result that holds throughout the period \hyperlink{slide_fig1}{\beamergotobutton{See Figure 1}} \vspace{0.1cm}
\item[-] Omitting the additive term substantially biases the iceberg component upwards (Table 1) \vspace{0.2cm}
\begin{itemize}
\footnotesize{
\item[$\ast$] The ad-valorem cost is reduced by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.1cm}
\item[$\ast$] From 5.8\% to 3.2\% in maritime transport (mean value over the period) \vspace{0.1cm}
\item[$\ast$] From 5.1\% to 2.5\% in Air transport }\vspace{0.1cm}
\end{itemize}
\end{itemize}
\item Opening the black box of transport costs \hyperlink{app_comparison}{\beamergotobutton{More on this}} \vspace{0.1cm}
\begin{itemize}
\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in Air \& Vessel resp.
\item[-] Additive cost: 1.8\% and 2.9\% of the export price \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Valuable insights for related papers more theoretically-oriented, in need for calibration

@@

@@

\begin{frame}[label = app_comparison]
\frametitle{More on the size of additive transport costs}
\begin{itemize}
\item Comparison with Irarrazabal et al. (2015) \vspace{0.1cm}
\begin{itemize}
\item[-] Irarrazabal et al. (2015): Estimate the trade costs ratio $t/\tau$ = 14\% of the export price $\widetilde{p}$ \vspace{0.1cm}
\item[-] In 2004, our estimates lead to $t/\tau = 1.8\% \times \widetilde{p}$ (Air), $t/\tau = 2.8\%\times \widetilde{p}$ (Ocean) \vspace{0.1cm}
\item[$\Rightarrow$] Much lower !
\end{itemize}
\item But, our work: Estimate transport costs, not trade costs \vspace{0.1cm}
\begin{itemize}
\item[-] According to AVW, transport costs $\simeq$ 15\% of trade costs \vspace{0.1cm}
\item[-] Plugging this into Irarrazabal et al.'s (2015) estimates yields a ratio $t/\tau$ of transport costs equal to $2.1\% \times \widetilde{p}$
\item[$\Rightarrow$] Our estimates amount to 62\% (Air) and 53\% (Vessel) of Irarrazabal \& al.'s estimates (of transport costs)\vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Additive transport costs represent between 1/2 and 2/3 of total additive trade costs
\end{itemize}
\hyperlink{slide_result1}{\beamergotobutton{Back to slide}}
\end{frame}

@@

\item Hummels (2007) obtains the fitted ad-valorem rate in one step (the LHS is the observed price gap) while we use a 2-step approach (extract the fitted TC from the \textit{estimated} unfitted ad-valorem rate) - but a slight difference \vspace{0.1cm}
\item Hummels (2007) does not purge its measure of the fitted ad-valorem rate by the country-sector fixed effect  - but only a scale effect \vspace{0.1cm}
\item We separate the country-product fixed effects in two components (country / product), due to the number of fixed effects to estimate in the non-linear setting - robustness analysis to this point (in progress) \vspace{0.1cm}
\item We differ in the weighting scheme to obtain the yearly value of the fitted TC measure.  \vspace{0.1cm}
\begin{itemize}
\footnotesize
\item[-] Hummels (2007): The unweighed average value over the $i,k$ dimension \vspace{0.1cm}
\item[$\neq$] We weight each flow by its relative value in total trade flows observed in 1974 \vspace{0.1cm}
\item[-] Not much of a difference provided the weighting scheme in value is constant over time \vspace{0.1cm}
\item[-] Results robust to the alternative weighting scheme (apply Hummels's (2007) method) 
\end{itemize}

>>>>

\begin{frame}[label = slide_variancedecomp]
\frametitle{Empirical specification (2)}
\textbf{The estimation strategy} \vspace{0.1cm}
\begin{itemize}
\item Make some simplifying assumptions to get the estimated equation \vspace{0.1cm}
\item About the specification of both additive and multiplicative costs  \vspace{0.1cm}
\begin{itemize}
\item[(1)] Separability between the origin country $i$ and the product $k$ dimensions (as in Irarrazabal et al., 2015)\vspace{0.1cm}
\item[(2)] All products $k$ in a 3-digit sector $s$ share the same structure of costs \hyperlink{app_variancedecomp}{\beamergotobutton{More on this}}\vspace{0.1cm}
\item[$\Leftrightarrow$] Write $t_{ik}$ and $\tau_{ik}$ as:
\footnotesize
\begin{equation}
 \tau_{ik} = \tau_{i} \times \tau_{s(k)}, \qquad t_{ik} = t_{i} + t_{s(k)} \label{eq:specifTC}
 \end{equation}
 \normalsize
\end{itemize}
\item About the specification of the error term \vspace{0.1cm}
\begin{itemize}
\item[-] Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} -1>0$, the error term should be always positive and multiplicative \vspace{0.1cm}
\item[$\Rightarrow$] The equation of interest becomes:
\footnotesize
$$\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{s(k)} -1+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})$$
\normalsize
\begin{itemize}
\item[-] With $\epsilon_{ik}$ following a normal law centered on 0
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[label = slide_result1]
\begin{itemize}
\item \textbf{International transport costs: A sizeable additive component}  \vspace{0.2cm}
\begin{itemize}
\item[-] 48.2\% of total costs in average for vessel, 42.3\% for air  \vspace{0.1cm}
\item[-] Omitting the additive term substantially biases the iceberg component upwards (Table 1) \vspace{0.2cm}
\begin{itemize}
\footnotesize{
\item[$\ast$] The ad-valorem cost is reduced by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.1cm}
\item[$\ast$] From 5.8\% to 3.2\% in maritime transport (mean value over the period) \vspace{0.1cm}
\item[$\ast$] From 5.1\% to 2.5\% in Air transport }\vspace{0.1cm}
\end{itemize}
\end{itemize}
\item[$\Rightarrow$] Additive costs, a sizeable and structural dimension of international transportation costs \vspace{0.1cm}
\item Opening the black box of transport costs \hyperlink{app_comparison}{\beamergotobutton{More on this}} \vspace{0.1cm}
\begin{itemize}
\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in Air \& Vessel resp.
\item[-] Additive cost: 1.8\% and 2.9\% of the export price \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Valuable insights for related papers more theoretically-oriented, in need for calibration

@@

\begin{frame}[label=app_variancedecomp]
\frametitle{Variance decomposition exercise}
\begin{figure}[htbp]
\caption{Variance decomposition (observed cif-fas price gap)}
\label{fig:decomp_variance}
\begin{center}
\includegraphics[width=8cm, height=6cm]{variance_decomposition.pdf}
%{\footnotesize {OECD data} }
\end{center}
\end{figure}
\hyperlink{slide_variancedecomp}{\beamergotobutton{Back to slide}}

@@

\begin{frame}[label = app_comparison]
\frametitle{More on the size of additive transport costs}
\begin{itemize}
\item Comparison with Irarrazabal et al. (2015) \vspace{0.1cm}
\begin{itemize}
\item[-] Irarrazabal et al. (2015): Estimate the trade costs ratio $t/\tau$ = 14\% of the export price $\widetilde{p}$ \vspace{0.1cm}
\item[-] In 2004, our estimates lead to $t/\tau = 1.8\% \times \widetilde{p}$ (Air), $t/\tau = 2.8\%\times \widetilde{p}$ (Ocean) \vspace{0.1cm}
\item[$\Rightarrow$] Much lower !
\end{itemize}
\item But, our work: Estimate transport costs, not trade costs \vspace{0.1cm}
\begin{itemize}
\item[-] According to AVW, transport costs $\simeq$ 15\% of trade costs \vspace{0.1cm}
\item[-] Plugging this into Irarrazabal et al.'s (2015) estimates yields a ratio $t/\tau$ of transport costs equal to $2.1\% \times \widetilde{p}$
\item[$\Rightarrow$] In line with our results\vspace{0.1cm}
\end{itemize}
\end{itemize}
\hyperlink{slide_result1}{\beamergotobutton{Back to slide}}
\end{frame}

@@

\item Hummels (2007) obtains the fitted ad-valorem rate in one step (the LHS is the observed price gap) while we use a 2-step approach (extract the fitted TC from the \textit{estimated} unfitted ad-valorem rate) - but a slight difference \vspace{0.1cm}
\item Hummels (2007) does not purge its measure of the fitted ad-valorem rate by the country-sector fixed effect  - but only a scale effect \vspace{0.1cm}
\item We separate the country-product fixed effects in two components (country / product), due to the number of fixed effects to estimate in the non-linear setting - robustness analysis to this point (in progress) \vspace{0.1cm}
\item We differ in the weighting scheme to obtain the yearly value of the fitted TC measure.  \vspace{0.1cm}
\begin{itemize}
\footnotesize
\item[-] Hummels (2007): The unweighed average value over the $i,k$ dimension \vspace{0.1cm}
\item[$\neq$] In estimating Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), we weight each flow by its relative value in total trade flows  \vspace{0.1cm}
\item[-] Not much of a difference provided the weighting scheme in value is constant over time \vspace{0.1cm}
\item[-] Results robust to the alternative weighting scheme (apply Hummels's (2007) method)
\end{itemize}
